---
title: "Bayesian Analysis of the Impact of Passive Voice on Missing Entities"
author: "Julian Frattini"
date: '2023-12-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(patchwork)

library(ggdag)

library(brms)
library(marginaleffects)
source("../util/model-eval.R")
```

This document contains a Bayesian analysis of the hypotheses investigated by Femmer et al.^[Femmer, H., Kučera, J., & Vetrò, A. (2014, September). On the impact of passive voice requirements on domain modelling. In Proceedings of the 8th ACM/IEEE International Symposium on Empirical Software Engineering and Measurement (pp. 1-4).].

## Data Loading

```{r data-loading}
source("../util/data-loading.R")
d <- load.data()

# print the data to ensure that all variables have the correct type
str(d)
```


## Bayesian Analysis

### Causal Assumptions

(TODO: DAG)

(TODO: adjustment sets)

### Formula

```{r formula}
formula <- (entities.missing | trials(entities.expected) ~ 
              1 + # standard success in the task of finding actors
              passive + # impact of the use of passive voice
              (1|RID) + # impact of the difficulty of the individual requirements
              (1|PID) + # impact of the skill of individual participants
              ExpREAca + # impact of the participants academic experience in RE
              ExpREInd) # impact of the participants industrial experience in RE

get_prior(formula, family=binomial, data=d)
```

### Priors

```{r priors}
priors <- c(prior(normal(-1.5, 1), class = Intercept),
            prior(normal(0, 1), class = b),
            prior(weibull(2, 1), class = sd, coef = Intercept, group = RID),
            prior(weibull(2, 1), class = sd, coef = Intercept, group = PID),
            prior(exponential(1), class = sd))
```

```{r model-prior}
m.prior <-
  brm(data = d, family = binomial, formula, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    #file = "fits/m.entities.prior"
  )
```

```{r prior-predictive-check}
ndraws <- 400

priorpc <- brms::pp_check(m.prior, type="bars", ndraws=ndraws)

priorpc
```

```{r model}
m <-
  brm(data = d, family = binomial, formula, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4,
    file = "fits/m.entities"
  )
```

```{r posterior-predictive-check}
postpc <- brms::pp_check(m, type="bars", ndraws=ndraws)

postpc
```

```{r summary}
summary(m)
```

```{r}
conditional_effects(m, effects="passive")
```

```{r}
conditional_effects(m, effects="ExpREInd")
```

```{r}
evaluate.model(m)
```