---
title: "Data Preparation"
author: "Julian Frattini"
date: '2023-12-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(patchwork)
```

```{r paths}
path_data <- "../data"
path_figures <- "../../figures/demographics"
```

This file prepares the data for the Bayesian data analysis. It loads all individual data sheets, joins them, cleans them up, and saves the processed file for easier access.

## Loading

We start by loading the data which is disclosed at https://doi.org/10.5281/zenodo.7499290. This Zenodo repository contains a file called `CorrectAnswers.pdf` with all the data obtained during the experiment (responses as well as demographic data). We manually extracted this data into four separate sheets.

```{r data-load}
participants <- read.csv(file=file.path(path_data, "participants.csv"))
experience <- read.csv(file=file.path(path_data, "experience.csv"))
requirements <- read.csv(file=file.path(path_data, "requirements.csv"))
responses <- read.csv(file=file.path(path_data, "responses.csv"))
```

We corrected two mistakes in the data:

1. The general experience points `Gxp` of participant `P1` should be 11, not 10.
2. The general experience points `Gxp` of participant `A10` should be 18, not 12.

We identified these mistakes by comparing the values as reported in the `CorrectAnswers.pdf` with the raw responses contained in the `Form Responses-Active.csv` and `Form Responses-Passive.csv`.

Next, we join all four sheets to one data frame.

```{r data-join}
d <- responses %>% 
  full_join(participants) %>% # add information about the participants
  full_join(experience) %>% # add information about the participants' experience
  full_join(requirements) # get information about the expected actors etc. of the requirement
```

This joined table contains one row per created domain model. With $n_p=15$ participants and $n_r=7$ requirements, we are observing $15\times7=105$ created domain models.

## Cleanup

Next, we clean up the joined data by performing the following operations:

1. Introducing an explicit `passive` variable that encodes whether a domain model was created based on a requirements specification using active or passive voice
2. Cutting off the variables containing the number of missing actors, entities, and associations at the number of expected actors, entities, and associations.

We performed the second step by counting the number of actors, entities, and associations that were *expected* in the sample solution according to `CorrectAnswers.pdf`. This is the maximum value for *missing* actors, entities, and associations in a domain model, since a domain model cannot miss more objects than there were objects expected.

```{r data-join}
d <- d %>%  mutate(
    passive = if_else(startsWith(PID, 'P'), TRUE, FALSE), # introduce a factor variable for passive voice

    actors.missing = if_else(MAct > EAct, EAct, MAct), # cutoff missing actors by expected actors
    actors.found = EAct-actors.missing, # inverse missed to found actors
    actors.expected = EAct, # spell out the variable
  
    associations.missing = if_else(MAsc > EAsc, EAsc, MAsc),
    associations.found = EAsc-associations.missing,
    associations.expected = EAsc,
    
    entities.missing = if_else(MEnt > EEnt, EEnt, MEnt),
    entities.found = EEnt-entities.missing,
    entities.expected = EEnt,
    
    REQuizPerformance = RExp)
```

Finally, we select only the relevant attributes of the resulting table.

```{r data-select}
d <- d %>% 
  select(
    PID, RID,
    Age, Program, REQuizPerformance,
    ExpProgAca, ExpProgInd, ExpSEAca, ExpSEInd, ExpREAca, ExpREInd,
    passive,
    actors.expected, actors.found, actors.missing,
    associations.expected, associations.found, associations.missing,
    entities.expected, entities.found, entities.missing)
```

## Data Export

Store the data to make it reusable in the analyses.

```{r export}
write_csv(d, file=file.path(path_data, "ipv-data.csv"))
```
